Large-scale harmonic patterns clustering

### 2010-03-01 Paper/Paper.tex
improved

The original goal of the \textit{Shingles} was retrieval, i.e. find very
similar song segments based on their harmonic pattern. Part of the
challenge was to handle tons of overlapping song segments represented
by a rather long feature vector (a serie of chroma features).
The LSH algorithm proved a great solution for performing a $k$-nearest
neighbors ($k$nn) search in such a space. However, the paradigm of $k$nn has
an intrinsec scaling limit.

@@

We propose to move the \textit{Shingle} idea into the larger clustering paradigm.
The goal is to find typical patterns that can be seen as cluster centroids.
Those typical patterns help us understand a music collection from an
harmonic point of view. Segments of songs can be described in regard to their
proximity to the different centroids, thus creating some notion of distance
among the song segments. These cluster of similar patterns can have many usages:
\begin{enumerate}
\item describing the music space
\item finding potential similar songs and/or covers
\item eliminating quickly very wrong matches in a query system
\end{enumerate}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
$10$K & $100$K & $500$K & $2$M  \\ \hline
$0$ & $1$ & $2$ & $3$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
% RESULTS:    avg.       mean dist
% 10K         
% 100K
% 500K
% 2M

>>>>

Working with large data amount of audio is a goal and a challenge at the
same time. One of the most common task is to find the nearest neighbor,
based on some distance, to a given segment of music. This can be viewed
as \textit{retrieval} as you are looking for some specific target
based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
designed for that purpose, and recent improvements let them scale to
considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
is a pattern of stacked chroma features. Similarity between
\textit{Shingles} coresspond to comparing harmonic patterns. For this
reason, they were used for cover or remix recognition \cite{Casey2007}.

That being said, nearest neighbor search is only one aspect of the
analysis of large-scale music data sets. Finding $k$ neighbors for a
particular query does not tell you much about the music space.
Clustering is a larger and more flexible framework where the goal is
to divide the space in meaningful blocks. The starting point is not
a query, but all the data set itself. In this work, we bring the
\textit{Shingle} idea into that framework in order to discover
meaningful and typical music harmonic patterns. The underlying assumption
is that the music space is articulated around a certain number of
frequent patterns.

Note that clustering and neighbor search is not totally unrelated.
In practise, we would hope that close neighbors live in one or two
close clusters, thus clustering can be a preprocess step. Also,
knowing all $k$ neighbors of each sample let you do graph clustering,
a subset of the clustering algorithms.

\iffalse % ignored section
The original goal of the \textit{Shingles} was retrieval, i.e. find very
similar song segments based on their harmonic pattern. Part of the
challenge was to handle tons of overlapping song segments represented
by a rather long feature vector (a serie of chroma features).
The LSH algorithm proved a great solution for performing a $k$-nearest
neighbors ($k$nn) search in such a space. However, the paradigm of $k$nn has
an intrinsec scaling limit.

@@

We propose to move the \textit{Shingle} idea into the larger clustering 
paradigm.
\fi % end of ignored section

The goal is to find typical patterns that can be seen as cluster centroids.
Those typical patterns help us understand a music collection from an
harmonic point of view. Segments of songs can be described in regard to their
proximity to the different centroids, thus creating some notion of distance
among the song segments. These cluster of similar patterns can have many usages:
\begin{enumerate}
\item describing the music space
\item finding potential similar songs and/or covers
\item eliminating quickly very wrong matches in a query system
\end{enumerate}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
$10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
$0.143$ & $$ & $$ & $$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
% RESULTS:    avg.       mean dist
% 10K         2.012      0.143
% 50K
% 100K
% 500K
% 2M
